EuPPollNet: A European database of plant-pollinator networks

### 2024-09-04 Manuscript/EuPPollNet_Ms.qmd
affiliations

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995]. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022. Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019]. This approach maximized the potential number of studies that could be incorporated in this database. The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*". To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).

@@

EuPPollNet offers the largest set of plant-pollinator studies and networks compiled to date at European level. The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records. While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).

This likely reflects the fact that most plant and pollinator species are rare and geographically restricted. For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions. Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level. This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003]. In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require numerous sampling events within and across habitats, particularly for plant-pollinator interactions.

>>>>

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995]. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022. Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019]. This approach maximized the potential number of studies that could be incorporated in this database. The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*". To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a floral visitor, referred to also as pollinators despite not evaluating pollination efficiency, and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).

@@

EuPPollNet offers the largest set of plant-pollinator studies and networks compiled to date at European level. The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records. While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species). However, although our study likely has higher pollinator coverage, as we did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted. For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions. Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level. This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003]. In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
